PyTracer: Automatically Profiling Numerical Instabilities in Python

### 2022-05-10 IEEEtran/paper-annotated.tex
Update for major revision

\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xcolor,colortbl}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue!50!black,citecolor=blue!75!black,filecolor=black,urlcolor=blue!75!black}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{makecell}
\usepackage{ulem} % provides sout
\usepackage[frozencache=true,cachedir=minted-cache]{minted} 
\usepackage[htt]{hyphenat}      
\usepackage{subcaption}
\usepackage{graphicx,import}
\usepackage{booktabs}
\usepackage{soul}

%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[10pt,journal,compsoc]{../sty/IEEEtran}

@@

\definecolor{lightred}{rgb}{1.0,0.6,0.6}
\definecolor{lightgreen}{rgb}{0.6,1.0,0.6}

\DeclareRobustCommand{\add}[1]{{\sethlcolor{lightgreen}\hl{#1}}}
\DeclareRobustCommand{\remove}[1]{{\sethlcolor{lightred}\hl{#1}}}

\DeclareRobustCommand{\addCite}[1]{{\sethlcolor{lightgreen}\hl{\mbox{#1}}}}
\DeclareRobustCommand{\removeCite}[1]{{\sethlcolor{lightred}\hl{\mbox{#1}}}}

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
    \begin{abstract}
        Numerical stability is a crucial requirement of reliable scientific
        computing. However, despite the pervasiveness of Python in data science,
        analyzing large Python programs remains challenging due to the lack of
        scalable numerical analysis tools available for this language. To fill
        this gap, we developed \pytracer, a profiler to quantify numerical
        instability in Python applications. \pytracer transparently instruments
        Python code to produce numerical traces and visualize them interactively
        in a Plotly dashboard. We designed \pytracer to be agnostic to numerical
        noise model , allowing for tool evaluation through Monte-Carlo
        Arithmetic, random rounding, random data perturbation, or structured
        noise for a particular application. We illustrate \pytracer's
        capabilities by testing the numerical stability of key functions in both
        SciPy and Scikit-learn, two dominant Python libraries for mathematical
        modeling. Through these evaluations, we demonstrate \pytracer as a
        scalable, automatic, and generic framework for numerical profiling in
        Python.
    \end{abstract}
}
% Note that keywords are not normally used for peerreview papers.
%% \begin{IEEEkeywords}
%% Computer Society, IEEE, IEEEtran, journal, \LaTeX, paper, template.
%% \end{IEEEkeywords}}

@@

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figure/workflow.pdf}
    \caption{Pytracer workflow
        % \gkmod{}{Note: replace "Execution" with "Multiple Independent Executions", and maybe move it to be between the gears and the pkl files.}
    }
    \label{fig:workflow}
\end{figure}

@@

Three  main  approaches  exist  to  detect  numerical  instabilities:
stochastic  arithmetic,  uncertainty  or  sensitivity analysis [23], and random
seed analysis [11].  \hl{PyTracer is agnostic to the noise model used and therefore
    works for all methods mentioned}.  For our use cases, we focus on stochastic
arithmetic since it does not make any assumption about the noise model, in
contrast to sensitivity analysis\add{.}

Stochastic arithmetic leverages randomness to estimate numerical instabilities
coming from the use of floating-point representations. The main idea is to treat
round-off errors as random variables and characterize them statistically. Monte
Carlo Arithmetic (MCA)~\cite{parker1997monte} uses this principle by introducing
two improvements: (i) a virtual precision parameter, allowing to simulate
reduced working precision, and (ii) different perturbation modes to introduce
perturbations on function inputs or output. MCA replaces each floating-point
number $x$ with a stochastic counterpart computed as:
\[
    inexact(x) =  x + \beta^{e_x - t}\xi
\]
where $\beta$ is the number base, $e_x$ is the magnitude of $x$, $t$ the virtual
precision and $\xi \in (-\frac{1}{2},\frac{1}{2})$ is a random uniform variable.
Virtual precision allows for the simulation of reduced working precision. MCA
can be applied in three modes: Random Rounding (RR), Precision Bounding (PB),
and full MCA, which apply function $inexact$ to the output, the inputs, or both,
respectively. While the RR mode is equivalent to stochastic rounding, the PB
mode can also identify catastrophic cancellations.

Stochastic arithmetic quantifies numerical error using the number of significant
digits $s$ estimated among the sampled values. A common formula to determine
this number from MCA samples is presented in~\cite{parker1997monte}:
\begin{equation}
    s = -\log_{\beta}{ \left| \dfrac{\sigma}{\mu} \right|} \label{eq:sig-digits}
\end{equation}
where $\mu$ and $\sigma$ are the sample mean and standard deviation of a
variable sampled with MCA.  
Sohier et al.~\cite{sohier2018confidence} recently provided a generalization of
this formula to include confidence intervals.

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figure/pytracer_layout.pdf}
    \caption{Pytracer visualization layout.}
    \label{fig:visu-layout}
\end{figure}

@@

\section{Numerical evaluations}

We evaluated \pytracer on two widely-used Python libraries:
SciPy~\cite{virtanen2020scipy}, the main library for scientific computing in
Python, and  scikit-learn~\cite{pedregosa2011scikit}, a reference
machine-learning framework that implements a wide range of state-of-the-art
algorithms. For each library, we executed the software examples provided in
their respective GitHub repositories. We traced each example five times, having
activated MCA in the Python interpreter \add{as well as} BLAS, LAPACK, NumPy, SciPy, and
scikit-learn. \add{Only instrumented the Python interpreter would lead to
    only perturbe pure Python floating-point operations and thus to miss
    operations in compiled codes (i.e. Cython and shared libraries).}
\hl{We repeated the experiment with two MCA modes, Random Rounding
    (RR) and Full MCA, resulting in 2 experiments for each example}. We used the
default virtual precision in Verificarlo, namely 24 bits for single-precision
floating-point values and 53 bits for double-precision values, which simulates
machine error. We used Python-3.8.5, NumPy-1.19.1, SciPy-1.5.4, and
scikit-learn-0.23.2.

\subsection{Results overview}

Table~\ref{tab:pytracer_results_summary} summarizes \pytracer's outcome on SciPy
and scikit-learn examples. Columns \textit{trace} and \textit{parse} report on
the tracing and aggregation steps with three possible outcomes: 1) green for
success, meaning that the execution ended well and no issues were raised by the
application, 2) red for numerical error in the application, meaning that the
execution raised a Python error before the end, and 3) orange for errors due to
the fuzzy noise injection that are not considered numerical errors, meaning that
the execution raised a Python error but due to a perturbation that should not
have been introduced (i.e. a floating-point value representing an integer).
Overall, the tracing and parsing of SciPy and scikit-learn examples in Random
Rounding mode was entirely successful for 32/40 examples and showed an average
precision loss of 8 bits. The remaining examples failed due to numerical errors
in the libraries, which we discuss hereafter. The instrumentation with Full MCA
is more invasive and reduced the number of successful executions to 12/40. Among
the 28/40 failed executions, one failed due to an actual numerical error
(\texttt{Bayesian Ridge Regression} example) and the other ones failed due to
perturbations that should not have been introduced.

Column \textit{sigbits} in Table~\ref{tab:pytracer_results_summary} summarizes
the precision for the main outputs of SciPy and scikit-learn examples. We
measured the precision in significant bits on the final primary example output
by taking the element-wise mean for non-scalar values and, if the example
produces many outputs, the average value among all of them. The numerical
quality across all examples varies from 0 to 52, with an average of 44
significant bits. For both modes, we note that only two examples failed due to a
numerical error (\texttt{face\_recognition} for RR and \texttt{Bayesian Ridge
    Regression} for Full MCA).

@@

    \small

@@

        \begin{tabular}{|lll|c|c|c|c|c|c|}

@@

            \multicolumn{3}{|c}{ \multirow{2}{*}{Application} }                 & \multicolumn{3}{|c|}{Random Rounding (RR)}                                & \multicolumn{3}{c|}{Full MCA}                                                                                      \\
            \cline{4-9}
                                                                                &                                                                           &                                                          & trace  & parse    & sigbits & trace  & parse  & sigbits \\

@@

            \multicolumn{1}{c|}{ \multirow{3}{2em}{ \rotatebox{90}{FFT} }}      & \discreteCosineRF                                                         & \valid                                                   & \valid & \s{52}   & \warn   & \cross & \s{47}           \\ 
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \fftOneDRf                                               & \valid & \valid   & 41      & \warn  & \cross & -       \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \fftTwoDRf                                               & \valid & \valid   & \s{52}  & \valid & \valid & \s{52}  \\
            \cline{2-9}

@@

                                                                                & \interOneDRf                                                              & \valid                                                   & \valid & 51       & \warn   & \cross & -                \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \multiRf                                                 & \valid & \valid   & \s{51}  & \warn  & \cross & \s{-}   \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \bsplineRf                                               & \valid & \valid   & 10      & \valid & \valid & 10      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \splineOneDRf                                            & \valid & \cross   & \s{52}  & \warn  & \cross & \s{52}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \splineTwoDRf                                            & \valid & \valid   & 10      & \warn  & \cross & -       \\
            \cline{2-9}

@@

                                                                                & \bfgsRf                                                                   & \valid                                                   & \valid & \s{48}   & \valid  & \valid & \s{46}           \\
            \cline{3-9}
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \globalRf                                                & \valid & \cross   & 17      & \warn  & \cross & -       \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | SHGO\footnote{Simplicial Homology Global Optimization} & \na    & \na      & \s{25}  & \na    & \na    & \s{-}   \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | Dual Annealing                                         & \na    & \na      & 4       & \na    & \na    & -       \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | Differential Evolution                                 & \na    & \na      & \s{0}   & \na    & \na    & \s{-}   \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | Basin Hopping                                          & \na    & \na      & 40      & \na    & \na    & -       \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | SHGO Sobol                                             & \na    & \na      & \s{45}  & \na    & \na    & \s{-}   \\
            \cline{3-9}
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \slsqpRf                                                 & \valid & \valid   & 45      & \valid & \valid & 44      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \lsmRf                                                   & \valid & \valid   & \s{48}  & \warn  & \cross & \s{-}   \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \nelderRf                                                & \valid & \valid   & 44      & \valid & \valid & 42      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \ncgRf                                                   & \valid & \valid   & \s{48}  & \valid & \cross & \s{47}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \rootRf                                                  & \valid & \cross   & 52      & \warn  & \cross & 50      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \rootLargRf                                              & \valid & \cross   & \s{29}  & \valid & \cross & \s{26}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \rootLargePredRf                                         & \valid & \cross   & 42      & \valid & \cross & 42      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \trustExactRf                                            & \valid & \valid   & \s{48}  & \valid & \valid & \s{46}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \trustTruncRf                                            & \valid & \valid   & 48      & \valid & \valid & 46      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \trustNCGRf                                              & \valid & \valid   & \s{52}  & \valid & \valid & \s{47}  \\

@@

                                                                                & \AdaboostRf                                                               & \valid                                                   & \valid & 48       & \warn   & \cross & -                \\
            \cline{3-9}
            \multicolumn{2}{|c|}{}                                              & \brrRf                                                                    & \valid                                                   & \valid & \s{25}   & \cross  & \cross & \s{-}            \\
            \multicolumn{2}{|c|}{}                                              & | $1^{st}$ dataset                                                        & \na                                                      & \na    & 25       & \na     & \na    & -                \\
            \multicolumn{2}{|c|}{}                                              & | $2^{nd}$ dataset                                                        & \na                                                      & \na    & \s{nan}  & \na     & \na    & \s{-}            \\
            \cline{3-9}
            \multicolumn{2}{|c|}{}                                              & \onlineClassifierComparisonRf                                             & \valid                                                   & \valid & 20       & \warn   & \cross & -                \\
            \multicolumn{2}{|c|}{}                                              & \kmeansRf                                                                 & \valid                                                   & \valid & \s{52}   & \valid  & \cross & \s{-}            \\
            \multicolumn{2}{|c|}{}                                              & \covarianceRf                                                             & \valid                                                   & \cross & 48       & \warn   & \cross & -                \\
            \multicolumn{2}{|c|}{}                                              & \decisionRf                                                               & \valid                                                   & \valid & \s{50}   & \valid  & \valid & \s{50}           \\
            \multicolumn{2}{|c|}{}                                              & \digitsRf                                                                 & \valid                                                   & \valid & 52       & \warn   & \cross & 52               \\
            \multicolumn{2}{|c|}{}                                              & \faceRf                                                                   & \cross                                                   & \valid & \s{-}    & \warn   & \cross & \s{-}            \\
            \multicolumn{2}{|c|}{}                                              & \penaltyRf                                                                & \valid                                                   & \valid & -        & \valid  & \valid & -                \\
            \multicolumn{2}{|c|}{}                                              & \lassoRf                                                                  & \valid                                                   & \valid & \s{48}   & \warn   & \cross & \s{47}           \\
            \multicolumn{2}{|c|}{}                                              & \hyperplaneRf                                                             & \valid                                                   & \valid & 3        & \valid  & \cross & -                \\
            \multicolumn{2}{|c|}{}                                              & \mnistRf                                                                  & \valid                                                   & \valid & \s{43}   & \warn   & \valid & \s{-}            \\
            \multicolumn{2}{|c|}{}                                              & \multitaskRf                                                              & \valid                                                   & \valid & 50       & \warn   & \valid & 48               \\
            \multicolumn{2}{|c|}{}                                              & \ompRf                                                                    & \valid                                                   & \valid & \s{52}   & \valid  & \valid & \s{52}           \\
            \multicolumn{2}{|c|}{}                                              & \pcaRf                                                                    & \valid                                                   & \valid & 48       & \valid  & \valid & 48               \\
            \cline{3-9}
            \multicolumn{2}{|c|}{}                                              & \robustRf                                                                 & \valid                                                   & \valid & \s{50}   & \warn   & \cross & \s{48}           \\
            \multicolumn{2}{|c|}{}                                              & | Linear                                                                  & \na                                                      & \na    & 48       & \na     & \na    & 46               \\
            \multicolumn{2}{|c|}{}                                              & | RANSAC                                                                  & \na                                                      & \na    & \s{50}   & \na     & \na    & \s{48}           \\
            \cline{3-9}
            \multicolumn{2}{|c|}{}                                              & \toyRf                                                                    & \valid                                                   & \valid & 35       & \warn   & \cross & -                \\
            \cline{3-9}
            \multicolumn{2}{|c|}{}                                              & \svmRf                                                                    & \valid                                                   & \valid & \s{43.5} & \valid  & \cross & \s{-}            \\
            \multicolumn{2}{|c|}{}                                              & | Non-weighted                                                            & \na                                                      & \na    & 43       & \na     & \na    & -                \\
            \multicolumn{2}{|c|}{}                                              & | Weighted                                                                & \na                                                      & \na    & \s{44}   & \na     & \na    & \s{-}            \\
            \cline{3-9}

>>>>

\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xcolor,colortbl}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue!50!black,citecolor=blue!75!black,filecolor=black,urlcolor=blue!75!black}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{makecell}
\usepackage{ulem} % provides sout
\usepackage[frozencache=true,cachedir=minted-cache]{minted} 
\usepackage[htt]{hyphenat}      
\usepackage{subcaption}
\usepackage{graphicx,import}
\usepackage{booktabs}
\usepackage{soul}
%\usepackage{cancel}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[10pt,journal,compsoc]{../sty/IEEEtran}

@@

\definecolor{lightred}{rgb}{0.6,0.0,0.0}
\definecolor{lightgreen}{rgb}{0.8,1.0,0.8}

\DeclareRobustCommand{\add}[1]{\sethlcolor{lightgreen}\hl{#1}}
\DeclareRobustCommand{\remove}[1]{\textcolor{lightred}{\sout{#1}}}

\DeclareRobustCommand{\addCite}[1]{\textcolor{lightgreen}{\mbox{#1}}}

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
    \begin{abstract}
        Numerical stability is a crucial requirement of reliable scientific
        computing. However, despite the pervasiveness of Python in data science,
        analyzing large Python programs remains challenging due to the lack of
        scalable numerical analysis tools available for this language. To fill
        this gap, we developed \pytracer, a profiler to quantify numerical
        instability in Python applications. \pytracer transparently instruments
        Python code to produce numerical traces and visualize them interactively
        in a Plotly dashboard. We designed \pytracer to be agnostic to numerical
        noise model , allowing for tool evaluation through Monte-Carlo
        Arithmetic, random rounding, random data perturbation, or structured
        noise for a particular application. We illustrate \pytracer's
        capabilities by testing the numerical stability of key functions in both
        SciPy and Scikit-learn, two dominant Python libraries for mathematical
        modeling. Through these evaluations, we demonstrate \pytracer as a
        scalable, automatic, and generic framework for numerical profiling in
        Python.
    \end{abstract}

    % Note that keywords are not normally used for peerreview papers.
    \begin{IEEEkeywords}
        Computer arithmetic, Quality analysis and evaluation, Visualization
        techniques and methodologies, Tracing,  Stochastic analysis
    \end{IEEEkeywords}
}

@@

\section{Philosophy of \pytracer}

\add{PyTracer follows a three steps approach:}
\begin{enumerate}
    \item \add{Automatic instrumentation of Python programs to save the value of inputs and outputs of function in a file.}
    \item \add{Aggregation of the traces generated to produce a statistics summary.}
    \item \add{Interactive visualization to navigate through the code execution.}
\end{enumerate}

\add{
    The main goal of PyTracer is to help the developper visualizing the numerical
    stability of Python applications. PyTracer saves time for the user by
    instrumenting and aggregating values. In the same way that GDB makes the user
    save time when debugging an application by providing all information in one
    layout, PyTracer allows the user to visualize the overall numerical quality to
    quickly detect instabilities. The PyTracer visualization layout improves result
    analysis in any case since 1) the user does not have to manually instrument its
    code to dump raw data and creates specific post-processing scripts to visualize
    it 2) PyTracer allows the user to export the raw data traced to extend the
    analysis. 
}

\add{
    The main challenge addresses by PyTracer is the transparent instrumentation of Python 
    applications. Indeed, since Python is interpreted language, one has no complete 
    view of the function that will be executed by a given program. 
    Moreover, since we target Python application used in HPC environments, one has
    most of the time no access to source codes or with limited write access.
    Therefore, PyTracer tackles this issue by instrumenting the function dynamically 
    thanks to the Python introspection.
}

\add{
    The based-trace approach used by PyTracer permits to limit the 
    overhead at runtime since the statistics computation is reported to the postprocess analysis.
    However, it limits the aggregation to deterministic executions,
    although control-flow analysis can help to merge divergent execution paths.
}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.75\linewidth]{figure/workflow.pdf}
    \caption{Pytracer workflow
        % \gkmod{}{Note: replace "Execution" with "Multiple Independent
        % Executions", and maybe move it to be between the gears and the pkl
        % files.}
    }
    \label{fig:workflow}
\end{figure*}

@@

Three  main  approaches  exist  to  detect  numerical  instabilities:
stochastic  arithmetic,  uncertainty  or  sensitivity analysis [23], and random
seed analysis [11].  \add{PyTracer is agnostic to the noise model used and therefore
    works for all methods mentioned}.  For our use cases, we focus on stochastic
arithmetic since it does not make any assumption about the noise model, in
contrast to sensitivity analysis\add{.}

Stochastic arithmetic leverages randomness to estimate numerical instabilities
coming from the use of floating-point representations. The main idea is to treat
round-off errors as random variables and characterize them statistically. Monte 
Carlo Arithmetic (MCA)~\cite{parker1997monte} uses this principle by introducing
two improvements: (i) a virtual precision parameter, allowing to simulate
reduced working precision, and (ii) different perturbation modes to introduce
perturbations on function inputs or output. MCA replaces each floating-point
number $x$ with a stochastic counterpart computed as:
\[
    inexact(x) =  x + \beta^{e_x - t}\xi
\]
where $\beta$ is the number base, $e_x$ is the magnitude of $x$, $t$ the virtual
precision and $\xi \in (-\frac{1}{2},\frac{1}{2})$ is a random uniform variable.
Virtual precision allows for the simulation of reduced working precision. MCA
can be applied in three modes \add{but we only used the} Random Rounding (RR) \add{in our
    experiments} \remove{, Precision Bounding (PB), and full MCA,} which apply
function $inexact$ to the output. \remove{, the inputs, or both, respectively.
    While the RR mode is equivalent to stochastic rounding, the PB mode can also
    identify catastrophic cancellations.} \add{Random Rounding simulates roundoff errors.}

Stochastic arithmetic quantifies numerical error using the number of significant
digits $s$ estimated among the sampled values. A common formula to determine
this number from MCA samples is presented in~\cite{parker1997monte}:
\begin{equation}
    s = -\log_{\beta}{ \left| \dfrac{\sigma}{\mu} \right|} \label{eq:sig-digits}
\end{equation}
where $\mu$ and $\sigma$ are the sample mean and standard deviation of a
variable sampled with MCA. \add{This can be seen as the ratio signal over noise.}
Sohier et al.~\cite{sohier2021confidence} recently provided a generalization of
this formula to include confidence intervals.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.75\linewidth]{figure/pytracer_layout.pdf}
    \caption{Pytracer visualization layout.}
    \label{fig:visu-layout}
\end{figure*}

@@

\section{Results}

We evaluated \pytracer on two widely-used Python libraries:
SciPy~\cite{virtanen2020scipy}, the main library for scientific computing in
Python, and  scikit-learn~\cite{pedregosa2011scikit}, a reference
machine-learning framework that implements a wide range of state-of-the-art
algorithms. For each library, we executed the software examples provided in
their respective GitHub repositories. We traced each example five times, having
activated MCA in the Python interpreter \add{as well as} BLAS, LAPACK, NumPy,
SciPy, and scikit-learn. \add{Only instrumented the Python interpreter would
    lead to only perturbe pure Python floating-point operations and thus to miss
    operations in compiled codes (i.e. Cython and shared libraries).} \add{We
    repeated the experiment with two MCA modes, Random Rounding (RR) and Full MCA,
    resulting in 2 experiments for each example}. We used the default virtual
precision in Verificarlo, namely 24 bits for single-precision floating-point
values and 53 bits for double-precision values, which simulates machine error.
We used Python-3.8.5, NumPy-1.19.1, SciPy-1.5.4, and scikit-learn-0.23.2.

\add{Monte Carlo Arithmetic samples the floating-point variability of a given
    program, and hence the more MCA repetitions we have, the more accurate will be
    the sampling distribution. In theory, an infinite number of MCA repetitions
    should give the exact distribution. While in practice low sample size (3$\simeq$30)
    can give a good estimate, the issue is to capture "rare events" that could make the
    computation strongly diverge from the mean. For now, determining the number of
    MCA samples to use is still an open research question. 
    Recent theoretical works}~\cite{sohier2021confidence}
\add{on the significant digits computation allow computing an associated
    confidence interval depending on the sample size and the underlying distribution of the outputs.}

% \subsection{Results overview}

% Table~\ref{tab:pytracer_results_summary} summarizes \pytracer's outcome on SciPy
% and scikit-learn examples. Columns \textit{trace} and \textit{parse} report on
% the tracing and aggregation steps with three possible outcomes: 1) green for
% success, meaning that the execution ended well and no issues were raised by the
% application, 2) red for numerical error in the application, meaning that the
% execution raised a Python error before the end, and 3) orange for errors due to
% the fuzzy noise injection that are not considered numerical errors, meaning that
% the execution raised a Python error but due to a perturbation that should not
% have been introduced (i.e. a floating-point value representing an integer).
% Overall, the tracing and parsing of SciPy and scikit-learn examples in Random
% Rounding mode was entirely successful for 32/40 examples and showed an average
% precision loss of 8 bits. The remaining examples failed due to numerical errors
% in the libraries, which we discuss hereafter. The instrumentation with Full MCA
% is more invasive and reduced the number of successful executions to 12/40. Among
% the 28/40 failed executions, one failed due to an actual numerical error
% (\texttt{Bayesian Ridge Regression} example) and the other ones failed due to
% perturbations that should not have been introduced.

% Column \textit{sigbits} in Table~\ref{tab:pytracer_results_summary} summarizes
% the precision for the main outputs of SciPy and scikit-learn examples. We
% measured the precision in significant bits on the final primary example output
% by taking the element-wise mean for non-scalar values and, if the example
% produces many outputs, the average value among all of them. The numerical
% quality across all examples varies from 0 to 52, with an average of 44
% significant bits. For both modes, we note that only two examples failed due to a
% numerical error (\texttt{face\_recognition} for RR and \texttt{Bayesian Ridge
%     Regression} for Full MCA).

@@

    \normalsize

@@

        \begin{tabular}{|lll|c|c|c|}

@@

            \multicolumn{3}{|c}{ \multirow{2}{*}{Application} }                 & \multicolumn{3}{|c|}{Random Rounding (RR)}                                                                                                                         \\
            \cline{4-6}
                                                                                &                                                                           &                                                          & trace  & parse    & sigbits \\

@@

            \multicolumn{1}{c|}{ \multirow{3}{2em}{ \rotatebox{90}{FFT} }}      & \discreteCosineRF                                                         & \valid                                                   & \valid & \s{52}             \\ 
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \fftOneDRf                                               & \valid & \valid   & 41      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \fftTwoDRf                                               & \valid & \valid   & \s{52}  \\
            \cline{2-6}

@@

                                                                                & \interOneDRf                                                              & \valid                                                   & \valid & 51                 \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \multiRf                                                 & \valid & \valid   & \s{51}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \bsplineRf                                               & \valid & \valid   & 10      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \splineOneDRf                                            & \valid & \cross   & \s{52}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \splineTwoDRf                                            & \valid & \valid   & 10      \\
            \cline{2-6}

@@

                                                                                & \bfgsRf                                                                   & \valid                                                   & \valid & \s{48}             \\
            \cline{3-6}
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \globalRf                                                & \valid & \cross   & 17      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | SHGO\footnote{Simplicial Homology Global Optimization} & \na    & \na      & \s{25}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | Dual Annealing                                         & \na    & \na      & 4       \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | Differential Evolution                                 & \na    & \na      & \s{0}   \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | Basin Hopping                                          & \na    & \na      & 40      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & | SHGO Sobol                                             & \na    & \na      & \s{45}  \\
            \cline{3-6}
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \slsqpRf                                                 & \valid & \valid   & 45      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \lsmRf                                                   & \valid & \valid   & \s{48}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \nelderRf                                                & \valid & \valid   & 44      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \ncgRf                                                   & \valid & \valid   & \s{48}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \rootRf                                                  & \valid & \cross   & 52      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \rootLargRf                                              & \valid & \cross   & \s{29}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \rootLargePredRf                                         & \valid & \cross   & 42      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \trustExactRf                                            & \valid & \valid   & \s{48}  \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \trustTruncRf                                            & \valid & \valid   & 48      \\
            \multicolumn{1}{|c|}{}                                              & \multicolumn{1}{c|}{}                                                     & \trustNCGRf                                              & \valid & \valid   & \s{52}  \\

@@

                                                                                & \AdaboostRf                                                               & \valid                                                   & \valid & 48                 \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}                                              & \brrRf                                                                    & \valid                                                   & \valid & \s{25}             \\
            \multicolumn{2}{|c|}{}                                              & | $1^{st}$ dataset                                                        & \na                                                      & \na    & 25                 \\
            \multicolumn{2}{|c|}{}                                              & | $2^{nd}$ dataset                                                        & \na                                                      & \na    & \s{nan}            \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}                                              & \onlineClassifierComparisonRf                                             & \valid                                                   & \valid & 20                 \\
            \multicolumn{2}{|c|}{}                                              & \kmeansRf                                                                 & \valid                                                   & \valid & \s{52}             \\
            \multicolumn{2}{|c|}{}                                              & \covarianceRf                                                             & \valid                                                   & \cross & 48                 \\
            \multicolumn{2}{|c|}{}                                              & \decisionRf                                                               & \valid                                                   & \valid & \s{50}             \\
            \multicolumn{2}{|c|}{}                                              & \digitsRf                                                                 & \valid                                                   & \valid & 52                 \\
            \multicolumn{2}{|c|}{}                                              & \faceRf                                                                   & \cross                                                   & \valid & \s{-}              \\
            \multicolumn{2}{|c|}{}                                              & \penaltyRf                                                                & \valid                                                   & \valid & -                  \\
            \multicolumn{2}{|c|}{}                                              & \lassoRf                                                                  & \valid                                                   & \valid & \s{48}             \\
            \multicolumn{2}{|c|}{}                                              & \hyperplaneRf                                                             & \valid                                                   & \valid & 3                  \\
            \multicolumn{2}{|c|}{}                                              & \mnistRf                                                                  & \valid                                                   & \valid & \s{43}             \\
            \multicolumn{2}{|c|}{}                                              & \multitaskRf                                                              & \valid                                                   & \valid & 50                 \\
            \multicolumn{2}{|c|}{}                                              & \ompRf                                                                    & \valid                                                   & \valid & \s{52}             \\
            \multicolumn{2}{|c|}{}                                              & \pcaRf                                                                    & \valid                                                   & \valid & 48                 \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}                                              & \robustRf                                                                 & \valid                                                   & \valid & \s{50}             \\
            \multicolumn{2}{|c|}{}                                              & | Linear                                                                  & \na                                                      & \na    & 48                 \\
            \multicolumn{2}{|c|}{}                                              & | RANSAC                                                                  & \na                                                      & \na    & \s{50}             \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}                                              & \toyRf                                                                    & \valid                                                   & \valid & 35                 \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}                                              & \svmRf                                                                    & \valid                                                   & \valid & \s{43.5}           \\
            \multicolumn{2}{|c|}{}                                              & | Non-weighted                                                            & \na                                                      & \na    & 43                 \\
            \multicolumn{2}{|c|}{}                                              & | Weighted                                                                & \na                                                      & \na    & \s{44}             \\
            \cline{3-6}
            \multicolumn{2}{|c|}{}                                              & \tomographyRf                                                             & \valid                                                   & \cross & -                  \\
            \multicolumn{2}{|c|}{}                                              & \weightedRf                                                               & \valid                                                   & \valid & \s{44}             \\
